### 2022-11-10 NSF2022/Description/backgroundv2.tex
BG almost done

(Surface-director models.)

@@

This section shows the principles of PDE-driven
self-organization apply to the concrete
problem of \emph{vesicle hydrodynamics}.
A vesicle is a small, fluid filled membrane sack.
Microscopically, the membrane of a vesicle is comprised
elongated amphiphilic lipids that form inner and outer monolayers.  
The goal in vesicle hydrodynamics is to study the flow patterns e.g.,
tank-treading, tumbling, that arrise due to the interaction
between membrane elasticity and surrounding fluid.
Traditional models represent a vesicle as a
continuous surface or curve $\Sigma$.
The elastic energy takes the form
\begin{equation}
\label{eq:Canham-Helfrich}
  \int_{\Sigma} k_B(H - k_0)^2\, dS 
\end{equation}
where $H$ is the mean curvature, $k_0$ is a spontaneous curvature,
and $k_B$ is a bending modulus. 
The variation of \eqref{eq:Canham-Helfrich} gives rise to a surface force
that pointwise balances the jump in hydrodynamic stress across the interace.
A vesicle is closed surface, and so while energy does involve
the Gaussian curvature, it does not enter \eqref{eq:Canham-Helfrich}
since Gauss-Bonnet Theorem guarantees that
the integral of Gaussian curvature is independent of the shape of $\Sigma$.

Depending on the situation,
one can further enforce constant volume constraints or permeability relationships.
If the interface is area incompressible, then the surface divergence
of velocity is zero.  On the other hand, if the surface is area compressible,
then \eqref{eq:Canham-Helfrich} contains an additional stretching term of the
form $k_A(\alpha - \alpha_0)^2/\alpha_0$ where $\alpha$ and $\alpha_0$ are the
deformed and equilibrium area densitities respectively.

\begin{wrapfigure}[12]{r}{0.5\textwidth}
  \vspace{-5pt}
\centerline{\includegraphics[width=0.5\textwidth]{figures/SA1Figures/AmphiphilicAssembly.pdf}}
  \vspace{-5pt}
\caption{\label{fig:amphiphilic_assembly}
Self-organization under amphiphilic boundary condition.
(a) Granules first match hydrophobic tails (red),
(b) then align length-wise.
The principle of self-organization holds in three dimensions 
as simulated by Kohl, Corona, and Veerapaneni
\cite{koh-cor-che-vee2021}.}
\end{wrapfigure}
There are downsides to traditional models in terms of generality and physical realism.
(i) Motivated by drug delivery applications, experimenters use electric
fields, for example, to controllably open and close pores in membranes
\cite{}.
These kinds of topological changes are clumsy to incorporate into 
sharp interface formulations that assume a continuous surface \cite{}.
(ii) Sharp-interface models assume a membrane of neglible thickness. 
However, in biological processes
like fusion, fission, or insertion and penetration by proteins and peptides, 
the length scales of the
deformation (tens of nanometers) are comparable to membrane thickness (about 5 nm).
Moreover, these processes crucially rely on variable lipid
orientation relative to the surface normal (the so-called \emph{tilt}
deformation),
whereas \eqref{eq:Canham-Helfrich} implicitly assumes that the lipids
are everywhere normal to the surface.  
(iii) Finally, realistic membranes are comprised of a mixtures of lipids and proteins.
Recent work \cite{} has incorporated lipid mixtures in \eqref{eq:Canham-Helfrich},
but these models are formidably complicated 
due to the additional constraints and transport equations involved.

Diffusive interface models and surface-director models partially make up for some of
these shortcomings. 
Topological changes are permitted in phase field and level set formulations,
but these methods smear out the details of lipid orientations.
The question lipid orientations is closely related to the vast and well-studied
field of nematic liquid crystals.  Here there has been recent development
in the simulation and theory of nematic liquid crystals on surfaces.
These models consider the interplay between defects and surface geometry \cite{},
finite element formulations \cite{}, and existence of minimizers \cite{}.

\subsubsection{Main modeling innovations}
We developed the model \eqref{eq:RBT}-\eqref{eqn:hydro_stress} 
as a robust and flexible approach
that cuts across many of the aforementioned challenging aspects to vesicle modeling.
(i) The granules are not fixed to a stencil, 
affording flexibility in terms of vesicle morphology and topology--sufficiently
large forces will pull the granules apart as physically required.
(ii) The model formulation is built around important details like lipid orientations,
and exhibits and inter-monolayer slip not present in diffusive interface models.
(iii) It is straightforward to consider mixtures of granules with different
types of shapes and boundary conditions of the granules.
With the help of boundary integral equations
and fast summation methods, the dynamics can be solved in 
near-linear complexity in $N_b$, which is equal to or better than 
the computational complexity of related sharp interface methods.

@@

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/Background/coarsening.pdf}
\end{center}
\vspace{-20pt}
\caption{\label{fig:coarsening}
Modeling procedure.
We initilize a suspension of granules.
Under the moving domain problem
\eqref{eq:RBT}-\eqref{eqn:hydro_stress}
with amphiphilic boundary condition, the granules
self-organize into vesicle bilayers.
We then introduce external hydrodynamic effects to study elastic properties.
}
\vspace{5pt}
\end{figure}
\subsubsection{How do granule-based vesicles have an elastic energy?}
A natural question to ask is how 
elastic energy enters a formulation
coming from surface interactions like \eqref{eqn:math_motivation}.
Conceptually, random suspensions of amphiphilic
granules undergo a rapid coarsening to form bilayers.
For small granule numbers, the bilayers form a single stable, ring-shaped vesicle
with energy $E_{\text{stable}}$ (Figure \ref{fig:coarsening}, (c)).
This vesicle configuration, however, is only a local minimizer,
whereas the planar configuration (Figure \ref{fig:coarsening}, (e))
is the global minimizer.  The difference between
$E(t)$ and $E_{\min}$ is the elastic energy stored by granules.
Indeed, if a background flow overcomes the energy barrier to force granules apart,
then bilayer unfolds from its circular shape into a flat shape.

When many granules are involved, the granules reach an intermediate
state consisting of several bilayer segments (see Figure \ref{fig:self-assembly2}, top right).
Placing this metastable state into a mild background flow
causes the segments to join up into isolated vesicles \cite{fu-ryh-qua-you2022}.

and we analyze its energy as follows:
\begin{enumerate}
\item[Bending]
  The vesicle has bending energy because the directors are not
  parallel.  In the outer layer, the directors have negative
  splay $\nabla \cdot \dd < 0$ and in the inner layer
  the directors have positive splay $\nabla \cdot \dd > 0$.
  Due to minimality, the energy depends locally quadratically
  on the splay and has a bending modulus $k_B$.
  When $\dd$ is everywhere parallel to the unit normal $\NN,$ then
  \[
  \nabla \dd = -2H
  \]
  where $H$ is the mean curvature. 
\item[Stretch]
  If the vesicle is stretched, then the distance between
  neighboring granules is increased from rest.  The energy
  for this increase is Hookean and for small deformations
  equals
  \begin{equation}
    \label{eq:stretch}
    \tfrac{1}{2}k_A(L - L_0)^2/L_0
  \end{equation}
  where $k_A$ is a stretching modulus, $L$ and $L_0$ are stretched
  vesicle length and length at rest, respectively
  ($\tfrac{1}{2}k_A(A - A_0)^2/A_0$ in three dimensions).
  If the stretching is too great, the vesicle ruptures
  and the free eneds either rejoin, repairing the vesicle,
  of move apart forming a flat bilayer.
\item[Tilt]
  The directors in Figure \ref{} are normal to the vesicle surface.
  However, there are situations where the directors are not normal
  to the surface and this introduces the tilt angle $\theta$
  given by $\cos \theta = \dd \cdot \NN$.
  The energy for tilt is also Hooken and behaves like
  $k_{\theta}|\dd \times \NN|^2$ for the tilt modulus $k_{\theta}$.
\end{enumerate}

\begin{wrapfigure}[11]{r}{0.5\textwidth}
\includegraphics[width=0.5\textwidth]{figures/PreliminaryWork/TankTreading.jpg}
\caption{\label{fig:JPv_linearshear}The JP vesicles undergo tank-treading in background shear flow.}
\end{wrapfigure}

@@

This is our first step to showing the energy of the granule-based
vesicles have the same elastic energy as bilayers in the limit $N_b \to \infty$.
Once we have simulations for large systems sizes, we can provide
evidence that the two formulations are the same.
So and so has made simulations with thousands of granules in three-D
and this is now a technical/not theoretical limitation.

>>>>

@@

This section shows the principles of PDE-driven
self-organization apply to the concrete
problem of \emph{vesicle hydrodynamics}.
A vesicle is a small, fluid filled membrane sack.
Microscopically, the membrane of a vesicle is comprised
elongated amphiphilic lipids that form inner and outer monolayers.  
The goal in vesicle hydrodynamics is to study the flow patterns e.g.,
tank-treading, tumbling, that arrise due to the interaction
between membrane elasticity and surrounding fluid.
Traditional models represent a vesicle as a
continuous surface (or curve in $\mathbb{R}^2$) $\Sigma$.
The elastic energy takes the form
\begin{equation}
\label{eq:Canham-Helfrich}
  \int_{\Sigma} k_B(H - k_0)^2\, dS 
\end{equation}
where $H$ is the mean curvature, $k_0$ is a spontaneous curvature,
and $k_B$ is a bending modulus. 
The variation of \eqref{eq:Canham-Helfrich} with respect to position
gives rise to a surface force
that pointwise balances the jump in hydrodynamic stress across the interace.
While energy density does depend on Gaussian curvature,
it does not appear in \eqref{eq:Canham-Helfrich}
since Gauss-Bonnet Theorem guarantees that
the integral of Gaussian curvature over a closed surface is independent of the shape of $\Sigma$.

Depending on the situation,
one can further enforce constant volume constraints or permeability relationships.
If the interface is area incompressible, then the surface divergence
of velocity is zero.  On the other hand, if the surface is area compressible,
then \eqref{eq:Canham-Helfrich} contains an additional stretching term of the
form $k_A(\alpha - \alpha_0)^2/\alpha_0^2$ where $\alpha$ and $\alpha_0$ are the
deformed and equilibrium area densitities respectively.

\begin{wrapfigure}[12]{r}{0.5\textwidth}
  \vspace{-5pt}
\centerline{\includegraphics[width=0.5\textwidth]{figures/SA1Figures/AmphiphilicAssembly.pdf}}
  \vspace{-5pt}
\caption{\label{fig:amphiphilic_assembly}
Self-organization under amphiphilic boundary condition.
(a) Granules first match hydrophobic tails (red),
(b) then align length-wise.
The principle of self-organization holds in three dimensions 
as simulated by Kohl, Corona, and Veerapaneni
\cite{koh-cor-che-vee2021}.}
\end{wrapfigure}
There are downsides to traditional models in terms of generality and physical realism.
(i) Motivated by drug delivery applications, experimenters use electric
fields, for example, to controllably open and close pores in membranes
\cite{}.
These kinds of topological changes are clumsy to incorporate in
sharp interface formulations that assume a continuous surface \cite{}.
(ii) Sharp-interface models assume a membrane of neglible thickness. 
However, in biological processes
like fusion, fission, or insertion and penetration by proteins and peptides, 
the length scales of the
deformation (tens of nanometers) are comparable to membrane thickness (about 5 nm).
Moreover, these processes crucially rely on variable lipid
orientation relative to the surface normal (the so-called \emph{tilt}
deformation),
whereas \eqref{eq:Canham-Helfrich} implicitly assumes that the lipids
are everywhere normal to the surface.  
(iii) Finally, realistic membranes are comprised of a mixtures of lipids and proteins.
Recent work \cite{} has incorporated lipid mixtures in \eqref{eq:Canham-Helfrich},
but these models are formidably complicated 
due to the additional constraints and transport equations involved.

Diffusive interface models and surface-director models partially address 
these shortcomings. 
Topological changes are permitted in phase field and level set formulations,
but these methods smear out the details of lipid orientations.
The detailed energies of lipid orientation that include tilt are
closely related to the celebraated and well-studied
field of nematic liquid crystals.  Here, there have been recent development
in simulation and theory of nematic liquid crystals on surfaces.
These models consider the interplay between defects and surface geometry \cite{},
finite element formulations \cite{}, and existence of minimizers \cite{}.

\subsubsection{Main modeling innovations}
We developed the model \eqref{eq:RBT}-\eqref{eqn:hydro_stress} 
as a robust and flexible approach
that cuts across many of the aforementioned challenges.
(i) The granules are not fixed to a stencil, 
affording flexibility in terms of vesicle morphology and topology--sufficiently
large forces will even pull the granules apart as physically required.
(ii) The model formulation is built around important details like lipid orientations,
and exhibits inter-monolayer slip that must be added into diffusive interface models
post hoc. 
(iii) It is straightforward to consider mixtures of granules with different
types of shapes and boundary conditions of the granules.
With the help of boundary integral equations
and fast summation methods, the dynamics can be solved in 
near-linear complexity in $N_b$, which is equal to or better than 
the computational complexity of related sharp interface methods.

@@

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/Background/coarsening.pdf}
\end{center}
\vspace{-20pt}
\caption{\label{fig:coarsening}
Modeling procedure.
We initilize a suspension of granules.
Under the moving domain problem
\eqref{eq:RBT}-\eqref{eqn:hydro_stress}
with amphiphilic boundary condition, the granules
self-organize into vesicle bilayers.
We then introduce external hydrodynamic effects to study elastic properties.
}
\vspace{5pt}
\end{figure}
\subsubsection{How do granule-based vesicles have an elastic energy?}
A natural question to ask is how 
elastic energy enters a formulation
coming from surface interactions like \eqref{eqn:math_motivation}.
Conceptually, random suspensions of amphiphilic
granules undergo a rapid coarsening to form bilayers.
For tens of granule, the bilayers form a single stable, ring-shaped vesicle
with energy $E_{\text{stable}}$ (Figure \ref{fig:coarsening}, (c)).
This vesicle configuration, however, is only a local minimizer,
whereas the planar configuration (Figure \ref{fig:coarsening}, (e))
is the global minimizer.  The difference between
$E(t)$ and $E_{\min}$ is the vesicle elastic energy.
Indeed, if a background flow overcomes the energy barrier to force granules apart,
then the bilayer unfolds from its circular shape into a flat shape.

When hundreds granules are involved,
the two-dimensional granules reach an intermediate
state consisting of several bilayer segments (see Figure \ref{fig:self-assembly2}, top right).
Placing this metastable state into a mild background flow
causes the segments to join up into isolated vesicles \cite{fu-ryh-qua-you2022}.

Having a non-specific model that replicates detailed
physics and biology is the crown jewel of applied mathematics
and we believe that ours is such a model.

\begin{wrapfigure}[11]{r}{0.5\textwidth}
\includegraphics[width=0.5\textwidth]{figures/PreliminaryWork/TankTreading.jpg}
\caption{\label{fig:JPv_linearshear}The JP vesicles undergo tank-treading in background shear flow.}
\end{wrapfigure}
\textit{Summary of results from the paper \cite{Fu2018_SIAM}.}
A striking aspect of the
model is its fidelity to physics.
The model contains a few
parameters; granule radius $c$,
screening length $\epsilon$,
surface tension $\gamma$.
When set to realistic values,
\cite{Fu2018_SIAM, ErLjCl89, Lin2005, Parsegian, Israelachvili80, GarciaSaez, KUZMIN2005, Petelska2012,Jackson2016},
we recover many of the well-known
elastic moduli for single-component lipid bilayers
\cite{Nagle17, Nagle17-2, LeVeWa14,NAGLE2000159}.
\emph{Bending:}
The bending deformation enters through the splay of the
granule directors.  In Figure \ref{fig:JPv_linearshear},
for example, the directors have negative, respectively positive,
splay $\nabla_{\Sigma} \cdot \dd$
in the outer, respectively inner monolayers.
Here $\nabla_{\Sigma}\cdot{} = \nabla \cdot {} - \mathbf{N}\mathbf{N}^T \nabla$
is the surface divergence for the unit normal $\NN$
and the splay energy density
for the granule configuration is Hookean with bending modulus $k_B$.
It is a fact from differential geometry that
when $\dd$ is everywhere parallel to $\NN,$
then $\nabla\cdot \dd = \pm 2H$ and this is how the
curvature term arises for \eqref{eq:Canham-Helfrich}
($2H = \kappa$ in the two-dimensional setting).
In the case of mixtures of granules of different sizes,
spontaneous curvature $k_0$ arises
from asymmetry in the monolayer leaflets.
\emph{Stretching.}
When stretched, the distance between
neighboring granules is increased from rest
and the energy for stretching behaves as 
$\tfrac{1}{2}k_A(L - L_0)^2/L_0$
for a stretching modulus $k_A$,
length $L$, and
length at rest $L_0$ of the two-dimensional vesicle. 
\emph{Tilt.}
Finally, the granules possess a 
tilt decay length scale $\lambda = \sqrt{k_B/k_{\theta}}$
well-known from the biophysical literature
for the case when the directors $\mathbf{d}$ are no longer normal
to the surface
\cite{KUZMIN2005}.

\textit{Summary of results from the paper \cite{FuQuRyYo22}.}

@@
